Paper Plots V2

### 2025-03-07 Rmd/paper_plots/paper_plots_1.3.Rmd
add scenarios

@@

# par(mfrow = c(4,1), mar = c(0,0,0,0), oma = c(2, 4, 2, .5), mgp = c(2, .6, 0))
# par(mfrow = c(3,1), mar = c(1.5, 3.1, 0, .5), mgp = c(2, .6, 0))
par(mfrow = c(2,1), mar = c(0,3.1,0,0), oma = c(2, 0, 2, .5), mgp = c(2, .6, 0))
plot(out_1.4$length$iss$YEAR, out_1.4$length$iss$iss, type = "l", ylab = "ISS", xaxt = "n", ylim = c(0, 15000), cex.lab = 1.3) 
lines(assessment_2023_iss$year, assessment_2023_iss$iss, col = "blue")
lines(out_1.4$length$iss$YEAR, out_1.4$length$iss$nhls, col = "red")
legend(1977, 11000, lty = c(1,1,1), col = c("black", "blue", "red"), legend = c("ISSB", "ISSA", "number of hauls"), bty = "n")
text(1977, 14000, "A")

@@

@@

>>>>

out_3.2 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_1.4_ISS_500iters_output_2025-03-06.RDS") ### mislabeled

@@

# par(mfrow = c(4,1), mar = c(0,0,0,0), oma = c(2, 4, 2, .5), mgp = c(2, .6, 0))
# par(mfrow = c(3,1), mar = c(1.5, 3.1, 0, .5), mgp = c(2, .6, 0))
par(mfrow = c(2,1), mar = c(0,3.1,0,0), oma = c(2, 0, 2, .5), mgp = c(2, .6, 0))
plot(out_1.4$length$iss$YEAR, out_1.4$length$iss$iss, type = "l", ylab = "ISS", xaxt = "n", ylim = c(0, 15000), cex.lab = 1.3) 
lines(assessment_2023_iss$year, assessment_2023_iss$iss, col = "blue")
lines(out_1.4$length$iss$YEAR, out_1.4$length$iss$nhls, col = "red")
lines(out_3.2$length$iss$YEAR, out_3.2$length$iss$iss, col = "green")
legend(1977, 11000, lty = c(1,1,1,1), col = c("black", "blue", "red", "green"), legend = c("ISSB", "ISSA", "number of hauls", "ISSB-noSS"), bty = "n")
text(1977, 14000, "A")

@@

### * Time series Expansion NEW
```{r, fig.width=9, fig.height=5}
out_2.1.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_30.1_ISS_500_iters_2025-03-01.RDS")

year = out_2.1.1$expansion_NULL$length$iss$YEAR

df = tidytable(year = year,
               none = out_2.1.1$expansion_NULL$length$iss$iss,
               medium = out_2.1.1$expansion_haul_numbers$length$iss$iss,
               full = out_2.1.1$expansion_haul_month_numbers$length$iss$iss)

# all years
ggplot(data = df2, aes(x = year, y = ISS, group = scenario, colour = scenario)) +
  geom_line() +
  labs(x = "", y = "ISSB") +
  geom_point() + 
  theme_classic() +
  theme(legend.position = c(0.84, 0.8),
        axis.title.x = element_blank())

# 1999:2023
df2 %>% 
  filter(year%in%1999:2023) -> df2
ggplot(data = df2, aes(x = year, y = ISS, group = scenario, colour = scenario)) +
  geom_line() +
  labs(x = "", y = "ISSB") +
  geom_point() + 
  theme_classic() +
  theme(legend.position = c(0.84, 0.8),
        axis.title.x = element_blank())
```

@@

### * DATA: NEW reduce samples
```{r}
out_31.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_31.1.TAS_ISS_500_iters_2025-03-02.RDS")

year = out_31.1$samples_1$data_H_P$length$iss$YEAR

### H_P ###################
df = tidytable(year = year,
               N1_HP = out_31.1$samples_1$data_H_P$length$iss$iss,
               N5_HP = out_31.1$samples_5$data_H_P$length$iss$iss,
               N10_HP = out_31.1$samples_10$data_H_P$length$iss$iss,
               N15_HP = out_31.1$samples_15$data_H_P$length$iss$iss,
               N20_HP = out_31.1$samples_20$data_H_P$length$iss$iss,

               NN_HP = out_1.4$length$iss$iss[out_1.4$length$iss$YEAR%in%1999:2023])

########## PROBLEM: these should be the same! I think the problem is that I ran it on TAS without updating baseISS from github.
# (out_27.1$length$iss$iss[out_27.1$length$iss$YEAR%in%1999:2023]-out_31.1$samples_20$data_H_P$length$iss$iss) %>% boxplot()
out_27.1$length$iss$iss[out_27.1$length$iss$YEAR%in%1999:2023]-out_31.1$samples_5$data_H_P$length$iss$iss
# (df2[scenario=="full"]$ISS - out_1.4$length$iss$iss[out_1.4$length$iss$YEAR%in%1999:2023]) %>% boxplot()

scenario_names = c("N1_HP", "N5_HP", "N10_HP", "N15_HP", "N20_HP", "NN_HP")

## 1
df2 %>% 
  tidytable::mutate(ISS.error = 100*(ISS - df2$ISS[df2$scenario=="N20_HP"])/df2$ISS[df2$scenario=="N20_HP"],
                    type = "relative_20",
                    .by = scenario) -> df3.1
## 2
df2 %>% 
  tidytable::mutate(ISS.error = 100*(ISS - df2$ISS[df2$scenario=="NN_HP"])/df2$ISS[df2$scenario=="NN_HP"],
                    type = "relative_N",
                    .by = scenario) -> df3.2

df3.1 %>% 
  tidytable::bind_rows(df3.2) %>%
  tidytable::filter(scenario!="NN_HP") %>% 
  tidytable::mutate(scenario = str_sub(scenario, start = 2, end = -4)) %>%
  tidytable::mutate(scenario = forcats::fct_inorder(scenario)) -> df4
df4 -> df4.a

### H ###################
df = tidytable(year = year,
               N1_H = out_31.1$samples_1$data_H$length$iss$iss,
               N5_H = out_31.1$samples_5$data_H$length$iss$iss,
               N10_H = out_31.1$samples_10$data_H$length$iss$iss,
               N15_H = out_31.1$samples_15$data_H$length$iss$iss,
               N20_H = out_31.1$samples_20$data_H$length$iss$iss,

               NN_H = out_1.4$length$iss$iss[out_1.4$length$iss$YEAR%in%1999:2023])

scenario_names = c("N1_H", "N5_H", "N10_H", "N15_H", "N20_H", "NN_H")

## 1
df2 %>% 
  tidytable::mutate(ISS.error = 100*(ISS - df2$ISS[df2$scenario=="N20_H"])/df2$ISS[df2$scenario=="N20_H"],
                    type = "relative_20",
                    .by = scenario) -> df3.1
## 2
df2 %>% 
  tidytable::mutate(ISS.error = 100*(ISS - df2$ISS[df2$scenario=="NN_H"])/df2$ISS[df2$scenario=="NN_H"],
                    type = "relative_N",
                    .by = scenario) -> df3.2

df3.1 %>% 
  tidytable::bind_rows(df3.2) %>%
  tidytable::filter(scenario!="NN_H") %>% 
  tidytable::mutate(scenario = str_sub(scenario, start = 2, end = -3)) %>%
  tidytable::mutate(scenario = forcats::fct_inorder(scenario)) -> df4
df4 -> df4.b

### H 20 only ###################
out_32.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_32.1_ISS_500_iters_2025-03-03.RDS")

year = out_32.1$samples_1$data_H_20$length$iss$YEAR

df = tidytable(year = year,
               N1_H20 = out_32.1$samples_1$data_H_20$length$iss$iss,
               N5_H20 = out_32.1$samples_5$data_H_20$length$iss$iss,
               N10_H20 = out_32.1$samples_10$data_H_20$length$iss$iss,
               N15_H20 = out_32.1$samples_15$data_H_20$length$iss$iss,
               N20_H20 = out_32.1$samples_20$data_H_20$length$iss$iss,

               NN_HP = out_1.4$length$iss$iss[out_1.4$length$iss$YEAR%in%1999:2023])

## 1
df2 %>% 
  tidytable::mutate(ISS.error = 100*(ISS - df2$ISS[df2$scenario=="N20_H20"])/df2$ISS[df2$scenario=="N20_H20"],
                    type = "relative_H20",
                    .by = scenario) -> df3.1
## 2
df2 %>% 
  tidytable::mutate(ISS.error = 100*(ISS - df2$ISS[df2$scenario=="NN_HP"])/df2$ISS[df2$scenario=="NN_HP"],
                    type = "relative_N",
                    .by = scenario) -> df3.2

df3.1 %>% 
  tidytable::bind_rows(df3.2) %>%
  tidytable::filter(scenario!="NN_HP") %>% 
  tidytable::mutate(scenario = str_sub(scenario, start = 2, end = -5)) %>%
  tidytable::mutate(scenario = forcats::fct_inorder(scenario)) -> df4
df4 -> df4.c

# 
# ### Port only ###################
# out_36.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_36.1_ISS_500_iters_2025-03-06.RDS")
# 
# year = out_36.1$samples_1$data_P$length$iss$YEAR
# 
# # # manually insert a zero for samples=1 year 2021 because only one port sample in 2021
# # out_36.1$samples_1$data_P$length$iss$iss[out_36.1$samples_1$data_P$length$iss$YEAR==2021] = 0
# # 
# 
# 
# df = tidytable(year = year,
#                N1_P = out_36.1$samples_1$data_P$length$iss$iss,
#                N5_P = out_36.1$samples_5$data_P$length$iss$iss,
#                N10_P = out_36.1$samples_10$data_P$length$iss$iss,
#                N15_P = out_36.1$samples_15$data_P$length$iss$iss,
#                N20_P = out_36.1$samples_20$data_P$length$iss$iss,
#                
#                NN_P = out_36.1$samples_999$data_P$length$iss$iss)
# 
# 
# df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")
# 
# 
# 
# ## 1
# df2 %>% 
#   tidytable::mutate(ISS.error = 100*(ISS - df2$ISS[df2$scenario=="N20_H20"])/df2$ISS[df2$scenario=="N20_H20"],
#                     type = "relative_H20",
#                     .by = scenario) -> df3.1
# ## 2
# df2 %>% 
#   tidytable::mutate(ISS.error = 100*(ISS - df2$ISS[df2$scenario=="NN_HP"])/df2$ISS[df2$scenario=="NN_HP"],
#                     type = "relative_N",
#                     .by = scenario) -> df3.2
# 
# 
# df3.1 %>% 
#   tidytable::bind_rows(df3.2) %>%
#   tidytable::filter(scenario!="NN_HP") %>% 
#   tidytable::mutate(scenario = str_sub(scenario, start = 2, end = -5)) %>%
#   tidytable::mutate(scenario = forcats::fct_inorder(scenario)) -> df4
# df4 -> df4.c
#

```{r, fig.height=8}
### plots

p1 = df4.a %>%
  ggplot(aes(x = scenario, y = ISS.error)) +
  geom_boxplot(aes(fill = type)) +
  labs(x = "", y = "ISSB % Difference", tag = "A") +
  theme_classic() +
  theme(legend.position = c(0.7, 0.15), 
        legend.direction = "horizontal", 
        legend.box = "horizontal",
        plot.tag.position = c(0.1, 0.95),
        plot.tag = element_text(size = 14, face = "bold")) + 
  guides(fill = guide_legend(title = "haul and port data",
                             title.position = "top", 
                             title.hjust = 0.5,
                             label.position = "bottom"))  +
  geom_hline(yintercept = -10, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  # geom_hline(yintercept = 10, linetype = "dashed", color = "black") +
  scale_y_continuous(breaks = seq(-100, 120, by = 10), limits = c(-80, 10))

p2 = df4.b %>%
  ggplot(aes(x = scenario, y = ISS.error)) +
  geom_boxplot(aes(fill = type)) +
  labs(x = "Samples per Haul", y = "ISSB % Difference", tag = "B") +
  theme_classic() +
  theme(legend.position = c(0.7, 0.15), 
        legend.direction = "horizontal", 
        legend.box = "horizontal",
        plot.tag.position = c(0.1, 0.95),
        plot.tag = element_text(size = 14, face = "bold")) + 
  guides(fill = guide_legend(title = "haul only data",
                             title.position = "top", 
                             title.hjust = 0.5,
                             label.position = "bottom"))  +
  geom_hline(yintercept = -10, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  # geom_hline(yintercept = 10, linetype = "dashed", color = "black") +
  scale_y_continuous(breaks = seq(-100, 120, by = 10), limits = c(-80, 10))

p3 = df4.c %>%
  ggplot(aes(x = scenario, y = ISS.error)) +
  geom_boxplot(aes(fill = type)) +
  labs(x = "Samples per Haul", y = "ISSB % Difference", tag = "optional C") +
  theme_classic() +
  theme(legend.position = c(0.7, 0.15), 
        legend.direction = "horizontal", 
        legend.box = "horizontal",
        plot.tag.position = c(0.1, 0.95),
        plot.tag = element_text(size = 14, face = "bold")) + 
  guides(fill = guide_legend(title = "haul=20 only data",
                             title.position = "top", 
                             title.hjust = 0.5,
                             label.position = "bottom"))  +
  geom_hline(yintercept = -10, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  # geom_hline(yintercept = 10, linetype = "dashed", color = "black") +
  scale_y_continuous(breaks = seq(-100, 120, by = 10), limits = c(-80, 10))

p1 + plot_spacer() + p2 + plot_layout(ncol = 1,
                           heights = c(6, -.5, 6))

p3

for above. "the median percent difference in ISSB increased with increasing sample size using the full data set (A) and only the haul data (B) over the period 1999-2023. There was no appreciable change in the percent difference when only using the haul data. There was also very little difference comparing the percent difference relative to the baseline ISSB or relative to applying the sample reduction routine at a sampling rate of 20. This verifies that the results are comparable for the following scenarios: comparisons to the baseline or comparisons to just using hauls with a sampling rate that never exceeds 20. The most conservative scenario is assuming a future sampling where high sampling rates from port samples are not available and the haul sampling rate is allowed to be below 20 but not exceed it. Comparisons relative to this scenario are the most relevant, conservative, scenario available to evaluate the impact of reducing observer haul sampling rate on composition data precision for future years. If the sampling rate behavior continues on average at least as good as it has done for the period 1999:2023 (which it is scheduled to do), we can expect the loss of precision not to exceed these values. (Assuming fishing trip and haul characteristics are comparable on average, also likely expected).

### NEW Haul and TRIP Decrease
```{r, eval=F}
out_29.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_29.1_ISS_500_iters_2025-03-02.RDS")

year = out_29.1$proportion_0.1$type_trip$length$iss$YEAR

# Trip
df = tidytable(year = year,
               N_0.1 = out_29.1$proportion_0.1$type_trip$length$iss$iss,
               N_0.2 = out_29.1$proportion_0.2$type_trip$length$iss$iss,
               N_0.3 = out_29.1$proportion_0.3$type_trip$length$iss$iss,
               N_0.4 = out_29.1$proportion_0.4$type_trip$length$iss$iss,
               N_0.5 = out_29.1$proportion_0.5$type_trip$length$iss$iss,
               N_0.6 = out_29.1$proportion_0.6$type_trip$length$iss$iss,
               N_0.7 = out_29.1$proportion_0.7$type_trip$length$iss$iss,
               N_0.8 = out_29.1$proportion_0.8$type_trip$length$iss$iss,
               N_0.9 = out_29.1$proportion_0.9$type_trip$length$iss$iss)

df2 %>% 
  tidytable::mutate(ISS.error = (ISS - out_1.4$length$iss$iss)/out_1.4$length$iss$iss,
                    type = "trip",
                    .by = scenario) -> df3.1

# Haul
df = tidytable(year = year,
               N_0.1 = out_29.1$proportion_0.1$type_haul$length$iss$iss,
               N_0.2 = out_29.1$proportion_0.2$type_haul$length$iss$iss,
               N_0.3 = out_29.1$proportion_0.3$type_haul$length$iss$iss,
               N_0.4 = out_29.1$proportion_0.4$type_haul$length$iss$iss,
               N_0.5 = out_29.1$proportion_0.5$type_haul$length$iss$iss,
               N_0.6 = out_29.1$proportion_0.6$type_haul$length$iss$iss,
               N_0.7 = out_29.1$proportion_0.7$type_haul$length$iss$iss,
               N_0.8 = out_29.1$proportion_0.8$type_haul$length$iss$iss,
               N_0.9 = out_29.1$proportion_0.9$type_haul$length$iss$iss)

df2 %>% 
  tidytable::mutate(ISS.error = (ISS - out_1.4$length$iss$iss)/out_1.4$length$iss$iss,
                    type = "haul",
                    .by = scenario) -> df3.2

df4 %>% 
  ggplot(aes(x = scenario, y = ISS.error*100)) +
  geom_boxplot(aes(fill = type)) +
  labs(x = "Proportional Change", y = "ISSB % Difference") +
  theme_classic() +
  theme(legend.position = c(0.8, 0.2)) +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  geom_hline(yintercept = -10, linetype = "dashed", color = "black") +
  scale_y_continuous(breaks = seq(-100, 20, by = 10), limits = c(-100, 20)) +
  scale_x_discrete(labels = c("0.1", "0.2", "0.3", "0.4", "0.5", "0.6", "0.7", "0.8", "0.9"))

### NEW Haul and TRIP Decrease w/o replacement
```{r}
out_29.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_33.1.TAS_ISS_500_iters_2025-03-04.RDS")

year = out_29.1$proportion_0.1$type_trip$length$iss$YEAR

# Trip
df = tidytable(year = year,
               N_0.1 = out_29.1$proportion_0.1$type_trip$length$iss$iss,
               N_0.2 = out_29.1$proportion_0.2$type_trip$length$iss$iss,
               N_0.3 = out_29.1$proportion_0.3$type_trip$length$iss$iss,
               N_0.4 = out_29.1$proportion_0.4$type_trip$length$iss$iss,
               N_0.5 = out_29.1$proportion_0.5$type_trip$length$iss$iss,
               N_0.6 = out_29.1$proportion_0.6$type_trip$length$iss$iss,
               N_0.7 = out_29.1$proportion_0.7$type_trip$length$iss$iss,
               N_0.8 = out_29.1$proportion_0.8$type_trip$length$iss$iss,
               N_0.9 = out_29.1$proportion_0.9$type_trip$length$iss$iss)

df2 %>% 
  tidytable::mutate(ISS.error = (ISS - out_1.4$length$iss$iss)/out_1.4$length$iss$iss,
                    type = "trip",
                    .by = scenario) -> df3.1

# Haul
df = tidytable(year = year,
               N_0.1 = out_29.1$proportion_0.1$type_haul$length$iss$iss,
               N_0.2 = out_29.1$proportion_0.2$type_haul$length$iss$iss,
               N_0.3 = out_29.1$proportion_0.3$type_haul$length$iss$iss,
               N_0.4 = out_29.1$proportion_0.4$type_haul$length$iss$iss,
               N_0.5 = out_29.1$proportion_0.5$type_haul$length$iss$iss,
               N_0.6 = out_29.1$proportion_0.6$type_haul$length$iss$iss,
               N_0.7 = out_29.1$proportion_0.7$type_haul$length$iss$iss,
               N_0.8 = out_29.1$proportion_0.8$type_haul$length$iss$iss,
               N_0.9 = out_29.1$proportion_0.9$type_haul$length$iss$iss)

df2 %>% 
  tidytable::mutate(ISS.error = (ISS - out_1.4$length$iss$iss)/out_1.4$length$iss$iss,
                    type = "haul",
                    .by = scenario) -> df3.2

df4 %>% 
  ggplot(aes(x = scenario, y = ISS.error*100)) +
  geom_boxplot(aes(fill = type)) +
  labs(x = "Proportional Change", y = "ISSB % Difference") +
  theme_classic() +
  theme(legend.position = c(0.8, 0.2)) +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  geom_hline(yintercept = -10, linetype = "dashed", color = "black") +
  scale_y_continuous(breaks = seq(-100, 20, by = 10), limits = c(-100, 20)) +
  scale_x_discrete(labels = c("0.1", "0.2", "0.3", "0.4", "0.5", "0.6", "0.7", "0.8", "0.9"))
